---
title: "PA1_template.Rmd"
author: "Luis Valladares"
date: "4 de junio de 2018"
output: html_document
---

```{r Libraries}
library(knitr)
library(ggplot2)
library(data.table)
```

```{r SetOptions, echo=TRUE}
opts_chunk$set(echo = TRUE)
```

1. Code for reading in the dataset and/or processing the data
```{r Reading}
dt <- fread('activity.csv', header = T)
dt$date <- as.Date(dt$date, '%Y-%m-%d')
```

2. Histogram of the total number of steps taken each day
```{r Histogram Steps}
dt0 <- na.omit(dt)
g <- ggplot(dt0, aes(steps))
g1 <- g + geom_histogram(fill = 'yellow') + labs(x = 'Number of steps taken each day') + scale_y_continuous(breaks=seq(0, 12000, 3000))
g1
```


3. Mean and median number of steps taken each day
```{r Mean and median}
x <- mean(dt$steps, na.rm = T)
x
x1 <- median(dt$steps, na.rm = T)
x1
```

4. Time series plot of the average number of steps taken
```{r Time series steps}
dt1 <- dt0[, .(stepsAve = mean(steps)), by = date]
h <- ggplot(dt1, aes(date, stepsAve))
h1 <- h + geom_line(color = "red") + labs(y = 'Average number of steps') + labs(title = 'Time series plot')
h1
```


5. The 5-minute interval that, on average, contains the maximum number of steps
```{r max steps}
dt[which.max(steps), ]
```


6. Code to describe and show a strategy for imputing missing data
```{r Imputing missing data}
dt2 <- dt[, .(stepsN = ifelse(is.na(steps) == T, ifelse(is.na(mean(steps)) == T, 0, mean(steps)), as.double(steps)), steps, interval), by = date]
dt2
```

7. Histogram of the total number of steps taken each day after missing values are imputed
```{r Histogram StepsN}
k <- ggplot(dt2, aes(stepsN))
k1 <- k + geom_histogram(fill = 'green') + labs(x = 'Number of steps taken each day') + scale_y_continuous(breaks=seq(0, 15000, 3000))
k1
```

8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r Plot Interval vs Average number of steps}
dt3 <- dt2[,.(date, TypeDay = ifelse(weekdays(date) %in% c('sábado', 'domingo'), 'weekend', 'weekday'), stepsN, steps, interval)]
dt3$TypeDay <- as.factor(dt3$TypeDay) 

dt4 <- dt3[, .(stepsAveXInt = mean(stepsN)), by = .(interval, TypeDay)]

m <- ggplot(dt4, aes(interval, stepsAveXInt))
m1 <- m + geom_line(color = "blue") + facet_wrap(~TypeDay, nrow = 2, ncol = 1) + labs(y = 'Average number of steps')
m1
```
